---
title: "Lab 9"
author: "Vignesh Narayanaswamy"
date: "3/28/2020"
output: html_document
---
```{r}
library(tidyverse)
library(lubridate)
```

## Data from daily reports

```{r}
report_03_30_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-30-2020.csv"))
```

### COVID19-related deaths in each US state as of 03/30/2020

```{r}
report_03_30_2020 %>% 
    filter (Country_Region == "US") %>% 
    ggplot(aes(x = Deaths,  y = reorder(Province_State, Deaths))) + 
    geom_point() +
    ggtitle("COVID19-related deaths for each US State") +
    ylab("Country/Region") +
    xlab("Deaths")
```

### The top 20 countries with the most reported COVID-19 related recovery as of 03/30/2020

```{r}
report_03_30_2020 %>% group_by(Country_Region) %>% summarise(Recovered = sum(Recovered)) %>% arrange(desc(Recovered)) %>% slice(1:20) %>% ggplot(aes(x = Recovered,  y = reorder(Country_Region, Recovered))) + geom_point() + ggtitle("The 20 countries with the most reported COV19-related recovery") + ylab("Country/Region") + xlab("Recovered")
```

## Data from time series

```{r}
time_series_global_confirmed<-read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>% rename(Province_State = "Province/State", Country_Region = "Country/Region")
```

```{r}
time_series_global_confirmed_long <- time_series_global_confirmed %>% pivot_longer(-c(Province_State, Country_Region, Lat, Long), names_to = "Date", values_to = "Confirmed")
```
```
```{r}
time_series_global_recovered <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>% rename(Province_State = "Province/State", Country_Region = "Country/Region")
```

```{r}
time_series_global_recovered_long <- time_series_global_recovered %>% pivot_longer(-c(Province_State, Country_Region, Lat, Long), names_to = "Date", values_to = "Recovered")
```

```{r}
time_series_global_confirmed_long <- time_series_global_confirmed_long %>% unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
```

```{r}
time_series_global_recovered_long <- time_series_global_recovered_long %>% unite(Key, Province_State, Country_Region, Date, sep = ".") %>% select(Key, Recovered)
```

```{r}
time_series_long_joined <- full_join(time_series_global_confirmed_long, time_series_global_recovered_long, by = c("Key"))
```

```{r}
time_series_global_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>% rename(Province_State = "Province/State", Country_Region = "Country/Region")
```

```{r}
time_series_global_deaths_long <- time_series_global_deaths_long %>% pivot_longer(-c(Province_State, Country_Region, Lat, Long), names_to = "Date", values_to = "Deaths")
```

```{r}
time_series_global_deaths_long <- time_series_global_deaths_long %>% unite(Key, Province_State, Country_Region, Date, sep = ".") %>% select(Key, Deaths)
```

```{r}
time_series_long_joined <- full_join(time_series_long_joined, time_series_global_deaths_long, by = c("Key")) %>% select(-Key)
```

```{r}
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
```

### COVID19 confirmed cases in India from 01/22/2020 to 03/30/2020

```{r}
time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Recovered"), sum) %>% 
    filter (Country_Region == "India") %>% 
    ggplot(aes(x = Date,  y = Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("India Confirmed COVID-19 Cases")
```

### Comparing 5 different countries on COVID19-related confirmed cases, recovered and deaths
```{r}
time_series_long_joined_counts <- time_series_long_joined %>% pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date), names_to = "Report_Type", values_to = "Counts")
```

```{r}
time_series_long_joined_counts %>% 
    group_by(Country_Region, Report_Type, Date) %>% 
    summarise(Counts = sum(Counts)) %>% 
    filter (Country_Region %in% c("India", "US", "Korea, South", "Italy", "China")) %>% 
    ggplot(aes(x = Date,  y = log2(Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Cases") + facet_wrap(~Country_Region, ncol=2, scales="free_y")
```

## Visualizing data

```{r}
library(maps)
library(viridis)
```

```{r}
world <- map_data("world")
mybreaks <- c(1, 20, 100, 1000, 50000)
```

### COVID19-related deaths worldwide over time

```{r}
time_series_global_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
    rename(Province_State = "Province/State", Country_Region = "Country/Region") %>%
    pivot_longer(-c(Province_State, Country_Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))
```

```{r}
static <- ggplot(subset(time_series_global_deaths_long,Date %in% seq(min(Date),max(Date),14)),
            aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
            borders("world", colour = NA, fill = "grey90") +
            theme_bw() +
            geom_point(shape = 21, color='red', fill='red', alpha = 0.5) +
            labs(title = 'COVID-19 deaths',x = '', y = '',
                 size="Cases (x1000))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,ncol=1)
static
```

### COVID19-related recovery worldwide over time

```{r}
time_series_recovered <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
    rename(Province_State = "Province/State", Country_Region = "Country/Region") %>%
    pivot_longer(-c(Province_State, Country_Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))
```

```{r}
static2 <- ggplot(subset(time_series_recovered, Date %in% seq(min(Date),max(Date),14)),
            aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
            borders("world", colour = NA, fill = "grey90") +
            theme_bw() +
            geom_point(shape = 21, color='green', fill='green', alpha = 0.5) +
            labs(title = 'COVID-19 recovery',x = '', y = '',
                 size="Cases (x1000))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,ncol=1)
static2
```

